
%%%%%%%%% BIOGRAPHICAL SKETCH -- 2 pages


% Your Bio should be divided into the following sections
% (a) Professional Preparation (education):
% Undergrad, Major, Year
% Graduate, Major, Year
% Postdoc, Area, Years-Inclusive
% (b) Appointments:  most recent first.
% (c) Publications:  5 related to the proposal, and 5 "Other Significant Publications"
% (d) Synergistic Activities (math-enhancing activities that were not
% part of your main job description, like editorial boards and
% conference organizing - any Math-related volunteer work.
% these are often similar to Broader Impacts

% (e) Collaborators & Other Affiliations: (use the following sections)
% list in alphabetical order, and include current affiliations parenthetically
% Collaborators and Co-editors: past 48 months.  If none, write "none"
% Graduate Advisors and Postdoctoral sponsors: (your own, no matter how long ago)
% Thesis advisor and postgraduate scholar-sponsor:  those you have advised
% in the past 5 years.  
% Total number of graduate students advised: X (all time)
% Total number of postdoctoral scholars sponsored: Y (all time)

\required{Biographical Sketch: Bruce A. Desmarais}

\begin{enumerate}
\item[\textbf{(a)}] \textbf{Professional Preparation:}
\begin{enumerate}
\item[--] Eastern Connecticut State University, Economics and Public Policy, B.A.  (2002)
\item[--] University of North Carolina at Chapel Hill, Political Science, M.A.  (2008)
\item[--] University of North Carolina at Chapel Hill, Political Science, Ph.D. (2010)
\end{enumerate}
\textbf{\item[(b)] Appointments:}
\begin{enumerate}
\item[--] \emph{University of Massachusetts Amherst}
Assistant Professor, 2010 - Present 
\end{enumerate}

\textbf{\item[(c)] Publications}
\begin{enumerate}
\item[(i)] Publications Directly Related to the Proposed Project
\begin{enumerate}
\item[--] Cranmer, Skyler J. and Bruce A. Desmarais. 2011. ``Inferential Network Analysis with Exponential Random Graph Models.'' \emph{Political Analysis.} 19(1): 66-86.
\item[--] Desmarais, Bruce A. and Skyler J. Cranmer.  2012.``Statistical Inference for Valued-Edge Networks: The Generalized Exponential Random Graph Model" \emph{PLoS-ONE.} 7(1):e30136. 
\item[--] Desmarais, Bruce A. and Skyler J. Cranmer. 2012. ``Statistical Mechanics of Networks: Estimation and Uncertainty.'' \emph{Physica A} 391(4): 1865-1876.
\item[--] Desmarais, Bruce A. and Skyler J. Cranmer. 2012. ``Micro-Level Interpretation of Exponential  Random Graph Models with Application to Estuary Networks'' \emph{Policy Studies Journal.} 40(3): 402-434.
\item[--] Cranmer, Skyler J., Tobias Heinrich, and Bruce A. Desmarais. Accepted 2012. ``Reciprocity and the Structural Determinants of the International Sanctions Network.''  \emph{Social Networks.}
\end{enumerate}



\item[(iii)] Other Significant Publications
\begin{enumerate}
\item[--] Desmarais, Bruce A. 2012.``Lessons in Disguise: Multivariate Predictive Mistakes in Collective Choice Models.'' \emph{Public Choice}.  151(3--4): 719-737..
\item[--] Cranmer, Skyler J., Bruce A. Desmarais, and Elizabeth J. Menninga. 2012.  ``Complex Dependencies in the Alliance Network'' \emph{Conflict Management and  Peace Science} 23(3).
\item[--] Cranmer, Skyler J., Bruce A. Desmarais, and Justin H. Kirkland. 2012. ``Towards a Network Theory of Alliance Formation'' \emph{International Interactions}. 38(3): 295-324.
\item[--] Harden, Jeffrey J. and Bruce A. Desmarais. 2011. ``Linear Models with Outliers: Choosing Between Conditional Mean and Conditional Median Methods'' \emph{State Politics and Policy Quarterly}. 11(4): 371-389.
\item[--] Desmarais, Bruce A. and Jeffrey J. Harden. 2012. ``Comparing Partial Likelihood and Robust Estimation Methods for the Cox Regression Model'' \emph{Political Analysis}. 20(1): 113-135.
\end{enumerate}
\end{enumerate}

\textbf{\item[(d)] Synergistic Activities}
\begin{enumerate}
\item[--] Participated in the founding and administration of the Triangle Political Methodology Group (2009--2010) -- \texttt{http://www.unc.edu/depts/polisci/methods/}. 
\item[--] Co-organized an interdisciplinary speaker series in Computational Social Science at UMass Amherst (2010--Present) -- \texttt{http://cssi.umass.edu/seminars.html}.
\item[--] Editorial board member, {\em State Politics \& Policy Quarterly}, (2011--Present).
\item[--] Member of the fellowship committee for the 2012 Political Networks Conference.
\end{enumerate}
\textbf{\item[(e)] Collaborators and other Affiliations}
\begin{enumerate}
\item[(i)] Collaborators
\begin{enumerate}
\item[--] Skyler Cranmer, University of North Carolina at Chapel Hill
\item[--] Jeffrey J. Harden, University of North Carolina at Chapel Hill
\item[--] Hanna Wallach, University of Massachusetts Amherst
\item[--] Brian Schaffner, University of Massachusetts Amherst
\item[--] Vincent Moscardelli, University of Connecticut
\item[--] Tobias Heinrich, Rice University
\item[--] Allison Freeman, Center for Community Capital (UNC Chapel Hill)
\item[--] Elizabeth Menninga, University of North Carolina Chapel Hill
\item[--] Justin Kirkland, University of North Carolina, Chapel Hill
\item[--] Rachel Shorey, University of Massachusetts Amherst
\item[--] Stuart Benjamin, Duke University
\item[--] Peter Krafft,  University of Massachusetts Amherst
\end{enumerate}

\item[(ii)] Graduate and Post-Doctoral Advisors
\begin{enumerate}
\item[--] Thomas Carsey, University of North Carolina Chapel Hill
\item[--] Skyler Cranmer, University of North Carolina Chapel Hill
\item[--] James Stimson, University of North Carolina Chapel Hill
\item[--] Kevin McGuire, University of North Carolina Chapel Hill
\item[--] Isaac Unah, University of North Carolina Chapel Hill
\end{enumerate}

\item[(ii)] Thesis Advisor
\begin{enumerate}
\item[--] Rachel Shorey, UMass Amherst Computer Science M.S. Student
\item[--] Peter Krafft, UMass Amherst Computer Science M.S. Student
\item[--] James Aaron, UMass Amherst Political Science Ph.D. Student
\item[--] Michael Kowal, UMass Amherst Political Science Ph.D. Student
\end{enumerate}
Total number of graduate students advised: 4

\end{enumerate}
\end{enumerate}

